---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bushtucker

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/bushtucker)](https://CRAN.R-project.org/package=bushtucker)
[![pkgdown](https://github.com/chrisbrownlie/bushtucker/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/chrisbrownlie/bushtucker/actions/workflows/pkgdown.yaml)
<!-- badges: end -->

{bushtucker} is a data package which provides datasets about the
24 seasons of the UK TV show "I'm A Celebrity, Get Me Out Of Here", spanning
from 2002 to 2024 and broadcast on ITV.

The show follows a group of UK celebrities camping in an Australian rainforest 
(or Welsh castle for seasons 20-21, due to COVID) for several weeks, completing 
physical and mental endurance challenges to earn meals and treats. These challenges are 
called 'Bushtucker Trials', hence the name of the package.

More information on the format can be found on 
[the show's Wikipedia pages](https://en.wikipedia.org/wiki/I'm_a_Celebrity...Get_Me_Out_of_Here!_(British_TV_series)), 
which is also where the data in this package comes from.

## Installation

You can install the development version of bushtucker like so:

```{r setup, eval = FALSE}
# install.packages("pak")
pak::pkg_install("chrisbrownlie/bushtucker")

# OR

# install.packages("remotes")
remotes::install_github("chrisbrownlie/bushtucker")
```

## Datasets

There are 6 key datasets included in this package:

- `seasons`: one row for each of the 24 seasons, includes info on location, presenters and dates
- `contestants`: one row per contestant per season, includes info on occupation, final position and trial performance
- `results`: one row per contestant per vote, includes the season and date, whether the contestant was eliminated during that vote, as well as the share of the vote they received (where available)
- `trials`: one row per trial, includes the season, name, episode, which contestant(s) took part and how they performed
- `challenges`: one row per challenge (Celebrity Chest, Dingo Dollar, Castle Coin or Deals on Wheels), includes the season, episode, contestant(s) that took part, whether the campmates got the question correct and the prize they received
- `ratings`: one row per episode, includes the total viewership and weekly ranking of the episode

Below you can see a snapshot of each dataset.

```{r glimpse_data}
library(bushtucker)
library(pillar) # for glimpse()

seasons |>
  tail() |>
  glimpse()

contestants |>
  tail() |>
  glimpse()

results |>
  tail() |>
  glimpse()

trials |>
  tail() |>
  glimpse()

challenges |>
  tail() |>
  glimpse()

ratings |>
  tail() |>
  glimpse()
```
